On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-03-31 manuscripts/PNAS/chazthm_pnas.tex
sml edits

\begin{definition}\label{SpecialSupportSet}\label{FriedrichsDefinition}
For subspaces $V_1, \ldots, V_\ell \subseteq \mathbb{R}^m$, set $r := 1$ when $\ell = 1$ and define for $\ell \geq 2$:
\begin{align*}
r(\{V_i\}_{i=1}^\ell) := 1 - \left(1 -  \max \prod_{i=1}^{\ell-1} \sin^2  \theta \left(V_i, \cap_{j>i} V_j \right)  \right)^{1/2},
\end{align*} 
%
where the maximum is taken over all possible orderings\footnote{We modify the quantity in \cite{Deutsch12} in this way since the subspace ordering is irrelevant to our purpose.} of the $V_i$ and where the angle $\theta \in (0,\frac{\pi}{2}]$ is defined in terms of its cosine as \cite[Def.~9.4]{Deutsch12}:
\begin{align*}
\cos{\theta(U,W)} := \max\left\{ |\langle \mathbf{u}, \mathbf{w} \rangle|: \substack{ \mathbf{u} \in U \cap (U \cap W)^\perp, \ \|\mathbf{u}\|_2 \leq 1 \\ \mathbf{w} \in W \cap (U \cap W)^\perp, \  \|\mathbf{w}\|_2 \leq 1 } \right\}.
\end{align*}
\end{definition}
Note that $\theta \in (0,\frac{\pi}{2}]$ implies $0 < r \leq 1$. (We acknowledge the somewhat counter-intuitive property that $\theta =  \pi/2$ when $U = V$ and when $U \perp V$.)  %\textbf{Note:} we can also define $\theta$ by \cite[Lem.~9.5]{Deutsch12}:
%\begin{align*}
%\cos{\theta(U,W)} := \max\left\{ |\langle \mathbf{u}, \mathbf{w} \rangle|: \mathbf{u} \in U \cap (U \cap W)^\perp, \|\mathbf{u}\|_2 \leq 1, \mathbf{w} \in W \|\mathbf{w}\|_2 \leq 1  \right\}.
%\end{align*}

@@

%%% No |S| = k assumption %%%
We begin by showing that for every $S \in E$ there is some $S' \subseteq [m]$, $|S'| \leq k$ for which the distance $d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_S'\})$ is controlled by $\varepsilon$. Fix $S \in E$. Since there are $(k-1){m \choose k}+1$ vectors $\mathbf{x}_i$ supported on $S$, the pigeon-hole principle implies that there exists some $S' \in {[m] \choose k}$ and some set of $k$ indices $I(S)$ such that the supports of all $\mathbf{x}_i$ and $\mathbf{\hat x}_i$ with $i \in I(S)$ are contained in $S$ and $S'$, respectively.

Let $\mathbf{X}$ and $\mathbf{\hat{X}}$ be the $m \times N$ matrices with columns $\mathbf{x}_i$ and $\mathbf{\hat x}_i$, respectively. It follows from the general linear position of the $\mathbf{x}_i$ and the linear independence of every $k$ columns of $\mathbf{A}$ that $L_k(\mathbf{AX}_{J(S)}) > 0$; that is, the columns of the $n \times k$ matrix $\mathbf{AX}_{J(S)}$ are linearly independent and thus form a basis for $\text{\rm Span}\{\mathbf{A}_{S}\}$. Fixing $\mathbf{y} \in \text{\rm Span}\{\mathbf{A}_{S}\}$, there then exists a unique $\mathbf{c} = (c_1, \ldots, c_k) \in \mathbb{R}^k$ such that $\mathbf{y} = \mathbf{AX}_{J(S)}\mathbf{c}$. Letting \mbox{$\mathbf{\hat{y}} = \mathbf{B\hat{X}}_{J(S)}\mathbf{c}$}, which is in $\text{\rm Span}\{\mathbf{B}_{S'}\}$, we have:
\begin{align*}
\|\mathbf{y} - \mathbf{\hat{y}}\|_2 
&= \|\sum_{i=1}^k c_i(\mathbf{AX}_{J(S)} - \mathbf{B\hat{X}}_{J(S)})_i\|_2
\leq \varepsilon \sum_{i=1}^k |c_i| \\
&\leq \varepsilon \sqrt{k}  \|\mathbf{c}\|_2 
%\leq \frac{\varepsilon \sqrt{k}}{L(\mathbf{A}X_{J(S)})} \|\mathbf{A}X_{J(S)}\mathbf{c}\|_2 
= \frac{\varepsilon}{L_k(\mathbf{AX}_{J(S)})} \|\mathbf{y}\|_2,
\end{align*}
where the last equality is by definition of $L_k$. It follows that for all $S \in E$,
\begin{align}\label{rhs222}
d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{S'}\}) 
\leq \frac{\varepsilon}{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) }.
%\leq \frac{\varepsilon}{L(\mathbf{AX}_{J(S)})}.
%&< \left( \frac{ L_2(\mathbf{A}) }{ 2^{|E|} \max_{j \in [m]} \|\mathbf{A}_j\|_2 } \right) \frac{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) } { L(\mathbf{AX}_{J(S)}) } \min_{\substack{F \subseteq E}} r( \{ \text{\rm Span}\{\mathbf{A}_{S}\} \}_{S \in F}) \\
\end{align}
The result then follows by  Lem.~\ref{MainLemma}, with the map $\pi: E \to {[m] \choose k}$ defined by the association $S \mapsto S'$ above.
%\begin{align}
%\delta= \frac{\varepsilon}{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) } = \frac{ C \varepsilon }{ \tilde C } < \frac{L_2(\mathbf{A})}{\tilde C}
%\end{align}
\end{proof}

@@

Plugging this into \eqref{f} and rearranging yields \eqref{DTILeq} after substituting $r(\{V_i\}_{i=1}^k)$ for $1 - z$ (as $z$ may depend on the arbitrary ordering of the $V_i$).
\end{proof}

>>>>

\begin{definition}\label{SpecialSupportSet}\label{FriedrichsDefinition}
For subspaces $V_1, \ldots, V_\ell \subseteq \mathbb{R}^m$, set $r := 1$ when $\ell = 1$ and define for $\ell \geq 2$:
\begin{align*}
r(\{V_i\}_{i=1}^\ell) := 1 - \left(1 -  \max \prod_{i=1}^{\ell-1} \sin^2  \theta \left(V_i, \cap_{j>i} V_j \right)  \right)^{1/2},
\end{align*} 
%
where the maximum is taken over all orderings\footnote{We modify the quantity in \cite{Deutsch12} in this way since the subspace ordering is irrelevant to our purpose.} of the $V_i$ and the angle $\theta \in (0,\frac{\pi}{2}]$ is defined implicitly as \cite[Def.~9.4]{Deutsch12}:
\begin{align*}
\cos{\theta(U,W)} := \max\left\{ |\langle \mathbf{u}, \mathbf{w} \rangle|: \substack{ \mathbf{u} \in U \cap (U \cap W)^\perp, \ \|\mathbf{u}\|_2 \leq 1 \\ \mathbf{w} \in W \cap (U \cap W)^\perp, \  \|\mathbf{w}\|_2 \leq 1 } \right\}.
\end{align*}
\end{definition}
Note that $\theta \in (0,\frac{\pi}{2}]$ implies $0 < r \leq 1$. (We acknowledge the somewhat counter-intuitive property that $\theta =  \pi/2$ when $U = V$ and when $U \perp V$.)  %\textbf{Note:} we can also define $\theta$ by \cite[Lem.~9.5]{Deutsch12}:
%\begin{align*}
%\cos{\theta(U,W)} := \max\left\{ |\langle \mathbf{u}, \mathbf{w} \rangle|: \mathbf{u} \in U \cap (U \cap W)^\perp, \|\mathbf{u}\|_2 \leq 1, \mathbf{w} \in W \|\mathbf{w}\|_2 \leq 1  \right\}.
%\end{align*}

@@

%%% No |S| = k assumption %%%
We begin by showing that for every $S \in E$ there is some $S' \subseteq [m]$, $|S'| \leq k$, for which the distance $d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{S'}\})$ is controlled by $\varepsilon$. Fix $S \in E$. Since there are $(k-1){m \choose k}+1$ vectors $\mathbf{x}_i$ supported on $S$, the pigeon-hole principle implies that there exists some $S' \in {[m] \choose k}$ and some set of $k$ indices $I(S)$ such that the supports of all $\mathbf{x}_i$ and $\mathbf{\hat x}_i$ with $i \in I(S)$ are contained in $S$ and $S'$, respectively.

Let $\mathbf{X}$ and $\mathbf{\hat{X}}$ be the $m \times N$ matrices with columns $\mathbf{x}_i$ and $\mathbf{\hat x}_i$, respectively. It follows from the general linear position of the $\mathbf{x}_i$ and the linear independence of every $k$ columns of $\mathbf{A}$ that $L_k(\mathbf{AX}_{J(S)}) > 0$; that is, the columns of the $n \times k$ matrix $\mathbf{AX}_{J(S)}$ are linearly independent and thus form a basis for $\text{\rm Span}\{\mathbf{A}_{S}\}$. Fixing $\mathbf{y} \in \text{\rm Span}\{\mathbf{A}_{S}\}$, there then exists a unique $\mathbf{c} = (c_1, \ldots, c_k) \in \mathbb{R}^k$ such that $\mathbf{y} = \mathbf{AX}_{J(S)}\mathbf{c}$. Letting \mbox{$\mathbf{\hat{y}} = \mathbf{B\hat{X}}_{J(S)}\mathbf{c}$}, which is in $\text{\rm Span}\{\mathbf{B}_{S'}\}$, we have:
\begin{align*}
\|\mathbf{y} - \mathbf{\hat{y}}\|_2 
&= \|\sum_{i=1}^k c_i(\mathbf{AX}_{I(S)} - \mathbf{B\hat{X}}_{I(S)})_i\|_2
\leq \varepsilon \sum_{i=1}^k |c_i| \\
&\leq \varepsilon \sqrt{k}  \|\mathbf{c}\|_2 
%\leq \frac{\varepsilon \sqrt{k}}{L(\mathbf{A}X_{J(S)})} \|\mathbf{A}X_{J(S)}\mathbf{c}\|_2 
= \frac{\varepsilon}{L_k(\mathbf{AX}_{I(S)})} \|\mathbf{y}\|_2,
\end{align*}
where the last equality is by definition of $L_k$. It follows that for all $S \in E$,
\begin{align}\label{rhs222}
d(\text{\rm Span}\{\mathbf{A}_S\}, \text{\rm Span}\{\mathbf{B}_{S'}\}) 
\leq \frac{\varepsilon}{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) }.
%\leq \frac{\varepsilon}{L(\mathbf{AX}_{J(S)})}.
%&< \left( \frac{ L_2(\mathbf{A}) }{ 2^{|E|} \max_{j \in [m]} \|\mathbf{A}_j\|_2 } \right) \frac{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) } { L(\mathbf{AX}_{J(S)}) } \min_{\substack{F \subseteq E}} r( \{ \text{\rm Span}\{\mathbf{A}_{S}\} \}_{S \in F}) \\
\end{align}
The result then follows by  Lem.~\ref{MainLemma}, with the map $\pi: E \to {[m] \choose k}$ defined by the association $S \mapsto S'$ above.
%\begin{align}
%\delta= \frac{\varepsilon}{ \min_{S \in E} L_k(\mathbf{AX}_{I(S)}) } = \frac{ C \varepsilon }{ \tilde C } < \frac{L_2(\mathbf{A})}{\tilde C}
%\end{align}
\end{proof}

@@

Substituting this into \eqref{f} and rearranging yields \eqref{DTILeq} after substituting $r(\{V_i\}_{i=1}^k)$ for $1 - z$ (as $z$ may depend on the arbitrary ordering of the $V_i$).
\end{proof}
